Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-09 paper.tex
small edits and added put into Infrastructure

\newcommand{\symb}[1]{\makebox{\it #1}} 
\newcommand\ie{i.e.\!\,, }
\newcommand\rel{Re${\cal{L}}$}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
\newcommand\ttand{\mbox{{$\land$}}}
\newcommand\ttor{\mbox{{$\lor$}}}
\newcommand\ttcap{\mbox{{$\cap$}}}
\newcommand\ttcup{\mbox{{$\cup$}}}
\newcommand\ttfun{\mbox{{$\Rightarrow$}}}
\newcommand\ttmimp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttimp{\mbox{{$\longrightarrow$}}}
\newcommand\ttequiv{\mbox{{$\equiv$}}}
\newcommand\ttexists{\mbox{{$\exists$}}}
\newcommand\ttforall{\mbox{{$\forall$}}}
\newcommand\ttneg{\mbox{{$\neg$}}}
\newcommand\ttneq{\mbox{{$\neq$}}}
\newcommand\ttin{\mbox{{$\in$}}}
\newcommand\ttnin{\mbox{{$\notin$}}}
\newcommand\ttImp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttlam{\mbox{\( \lambda \)}}
\newcommand\tttimes{\mbox{\( \times \)}}
\newcommand\ttlbrack{\mbox{\(\llbracket\)}}
\newcommand\ttrbrack{\mbox{\( \rrbracket \)}}
\newcommand\noie{\textit{i.e.},\xspace}
\newcommand\eg{~\textit{e.g.},\xspace}
\newcommand\noeg{\textit{e.g.},\xspace}
\newcommand\ttatI{\mbox{\( @_G \)}}
\newcommand\ttto[1]{\mbox{{$\to^{#1}$}}}
\newcommand\ttleq{\mbox{{$\le$}}}
\newcommand\ttrelI{\mbox{{$\to_{i}$}}}
\newcommand\ttrelIstar{\mbox{{$\to^*$}}}
\newcommand\ttrel[1]{\mbox{{$\to_{#1}$}}}
\newcommand\ttrelstar[1]{\mbox{{$\to_{#1}^*$}}}
\newcommand\ttalpha{\mbox{{$\alpha$}}}
\newcommand\ttsupseteq{\mbox{{$\supseteq$}}}
\newcommand\ttupdownarrow{\mbox{{$\Updownarrow$}}}
\newcommand\ttbigcap{\mbox{{$\bigcap$}}}
\newcommand\tttau{\mbox{{$\tau$}}}
\newcommand\ttsubseteq{\mbox{{$\subseteq$}}}
\newcommand\ttf{\mbox{{$f$}}}
\newcommand\ttvdash{\mbox{{$\vdash$}}}
\newcommand\ttref[1]{\mbox{\(\sqsubseteq_{#1}\)}}
\newcommand\ttrefV{\mbox{{$\sqsubseteq_V$}}}
\newcommand{\ttcalN}[1]{\mbox{{${\mathcal{N}}_{\texttt{#1}}$}}} 
\newcommand\ttNatt{\mbox{{$\mathcal{N}$}}}
\newcommand\ttattand[1]{\mbox{{$\oplus_{\wedge}^{#1}$}}}
\newcommand\ttattor[1]{\mbox{{$\oplus_{\vee}^{#1}$}}}
\newcommand\ttsigma{\mbox{{$\sigma$}}}
\newcommand\ttmref[1]{\mbox{{$\sqsubseteq_{#1}$}}}
\newcommand\ttmeref{\ttmref{\mathcal{E}}}
\newcommand\ttsn{\mbox{{\texttt{s}$_0$}}}
\newcommand\ttsigmap{\mbox{{$\sigma'$}}}
\newcommand\ttecal{\mbox{$\mathcal{E}$}}
\newcommand\ttimg{\mbox{$\triangleleft$}}

@@

\section{Refinement}
\label{sec:ref}
%\TODO{This section is just copied over from arxive paper - adapt}
Intuitively, a refinement changes some aspect of the type of
the state, for example, replaces a data type by a richer datatype or
restricts the behaviour of the actors. The former is expressed directly 
by a mapping of datatypes, the latter is incorporated into the state
transition relation of the Kripke structure that corresponds to the 
transformed model.
%This relationship between the Kripke structure can be better understood
%visually (see Figure \ref{fig:modtrans})
In other words, we can encode a refinement within our framework
as a relation on Kripke structures that is parameterized additionally by
a function that maps the refined type to the abstract type.
The direction ``from refined to abstract'' of this type mapping may seem 
curiously counter-intuitive. However, the actual refinement is given by the 
refinement that uses this function as an input. The refinement
then refines an abstract to a more concrete system specification. 
The additional layer for the refinement 
%(the red box in Figure \ref{fig:theorystruc})
can be formalized in Isabelle as a
%second\footnote{The first refinement 
%relation in this framework is on attack trees summarized in Section \ref{sec:at}.} 
refinement relation 
$\sqsubseteq_{\mathcal{E}}$. 
The relation \texttt{mod\_trans} is typed as a relation over triples --
a function from a threefold Cartesian product to \texttt{bool}, the 
type containing true and false only.  
The type variables $\sigma$ and $\sigma'$ input to the type constructor 
\texttt{Kripke} represent the abstract state type and the concrete state type. 
Consequently, the middle element of the triples selected by the relation 
\texttt{mod\_trans} is a function of type $\sigma' \Rightarrow \sigma$ 
mapping elements of the refined state to the abstract state.
The expression in quotation marks after the type is again the
infix syntax in Isabelle that allows the definition of mathematical notation
instead of writing \texttt{mod\_trans} in prefix manner. This nicer infix
syntax is already used in the actual definition.
Finally, the arrow \texttt{\ttImp} is the implication of Isabelle's  meta-logic
while $\ttimp$ is the one of the {\it object} logic HOL. 
They are logically equivalent but of different
types: within a HOL formula $P$, for example, as below $\forall x. P \ttimp Q$, only the implication $\ttimp$
can be used.
\begin{ttbox}
 mod_trans ::  (\ttsigma Kripke \tttimes (\ttsigma' \ttfun \ttsigma) \tttimes \ttsigma' Kripke) 
               \ttfun bool                  ("_ \ttmref{(\_)} _")
  K \ttmeref K' \ttequiv \ttforall s' \ttin states K'. \ttforall s \ttin init K'. 
             s \ttrelstar{\sigma'} s' \ttimp \ttecal(s) \ttin init K 
             \ttand \ttecal(s) \ttrelstar{\sigma} \ttecal(s')
\end{ttbox}
The definition of \texttt{K \ttmeref\, K'} states that for any state $s$ 
of the refined Kripke structure that can be reached by the state transition
in zero or more steps from an initial state $s_0$ of the refined Kripke 
structure, the mapping ${\mathcal E}$ from the refined to the abstract 
model's state must preserve this reachability, i.e., the image of
$s_0$ must also be an initial state and from there the image of $s$
under ${\mathcal E}$ must be reached with $0$ or $n$ steps.

>>>>

\newcommand{\symb}[1]{\makebox{\it #1}} 
\newcommand\ie{i.e.\!\,, }
\newcommand\rel{Re${\cal{L}}$}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
\newcommand\ttand{\mbox{{$\land$}}}
\newcommand\ttor{\mbox{{$\lor$}}}
\newcommand\ttcap{\mbox{{$\cap$}}}
\newcommand\ttcup{\mbox{{$\cup$}}}
\newcommand\ttfun{\mbox{{$\Rightarrow$}}}
\newcommand\ttmimp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttimp{\mbox{{$\longrightarrow$}}}
\newcommand\ttequiv{\mbox{{$\equiv$}}}
\newcommand\ttexists{\mbox{{$\exists$}}}
\newcommand\ttforall{\mbox{{$\forall$}}}
\newcommand\ttneg{\mbox{{$\neg$}}}
\newcommand\ttneq{\mbox{{$\neq$}}}
\newcommand\ttin{\mbox{{$\in$}}}
\newcommand\ttnin{\mbox{{$\notin$}}}
\newcommand\ttImp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttlam{\mbox{\( \lambda \)}}
\newcommand\tttimes{\mbox{\( \times \)}}
\newcommand\ttlbrack{\mbox{\(\llbracket\)}}
\newcommand\ttrbrack{\mbox{\( \rrbracket \)}}
\newcommand\noie{\textit{i.e.},\xspace}
\newcommand\eg{~\textit{e.g.},\xspace}
\newcommand\noeg{\textit{e.g.},\xspace}
\newcommand\ttatI{\mbox{\( @_G \)}}
\newcommand\ttto[1]{\mbox{{$\to^{#1}$}}}
\newcommand\ttleq{\mbox{{$\le$}}}
\newcommand\ttrelI{\mbox{{$\to_{i}$}}}
\newcommand\ttrelIstar{\mbox{{$\to^*$}}}
\newcommand\ttrel[1]{\mbox{{$\to_{#1}$}}}
\newcommand\ttrelstar[1]{\mbox{{$\to_{#1}^*$}}}
\newcommand\ttalpha{\mbox{{$\alpha$}}}
\newcommand\ttsupseteq{\mbox{{$\supseteq$}}}
\newcommand\ttupdownarrow{\mbox{{$\Updownarrow$}}}
\newcommand\ttbigcap{\mbox{{$\bigcap$}}}
\newcommand\tttau{\mbox{{$\tau$}}}
\newcommand\ttsubseteq{\mbox{{$\subseteq$}}}
\newcommand\ttf{\mbox{{$f$}}}
\newcommand\ttvdash{\mbox{{$\vdash$}}}
\newcommand\ttref[1]{\mbox{\(\sqsubseteq_{#1}\)}}
\newcommand\ttrefV{\mbox{{$\sqsubseteq_V$}}}
\newcommand{\ttcalN}[1]{\mbox{{${\mathcal{N}}_{\texttt{#1}}$}}} 
\newcommand\ttNatt{\mbox{{$\mathcal{N}$}}}
\newcommand\ttattand[1]{\mbox{{$\oplus_{\wedge}^{#1}$}}}
\newcommand\ttattor[1]{\mbox{{$\oplus_{\vee}^{#1}$}}}
\newcommand\ttsigma{\mbox{{$\sigma$}}}
\newcommand\ttmref[1]{\mbox{{$\sqsubseteq_{#1}$}}}
\newcommand\ttmeref{\ttmref{\mathcal{E}}}
\newcommand\ttsn{\mbox{{\texttt{s}$_0$}}}
\newcommand\ttsigmap{\mbox{{$\sigma'$}}}
\newcommand\ttecal{\mbox{$\mathcal{E}$}}
%\newcommand\ttimg{\mbox{$\triangleleft$}}
\newcommand\ttimg{\mbox{\texttt{`}}}

@@

\section{Refinement}
\label{sec:ref}
%\TODO{This section is just copied over from arxive paper - adapt}
Intuitively, a refinement changes some aspect of the type of
the state, for example, replaces a data type by a richer datatype or
restricts the behaviour of the actors. The former is expressed directly 
by a mapping of datatypes, the latter is incorporated into the state
transition relation of the Kripke structure that corresponds to the 
transformed model.
%This relationship between the Kripke structure can be better understood
%visually (see Figure \ref{fig:modtrans})
In other words, we can encode a refinement within our framework
as a relation on Kripke structures that is parameterized additionally by
a function that maps the refined type to the abstract type.
The direction ``from refined to abstract'' of this type mapping may seem 
curiously counter-intuitive. However, the actual refinement is given by the 
refinement that uses this function as an input. The refinement
then refines an abstract to a more concrete system specification. 
The additional layer for the refinement 
%(the red box in Figure \ref{fig:theorystruc})
can be formalized in Isabelle as a
%second\footnote{The first refinement 
%relation in this framework is on attack trees summarized in Section \ref{sec:at}.} 
refinement relation 
$\sqsubseteq_{\mathcal{E}}$. 
The relation \texttt{mod\_trans} is typed as a relation over triples --
a function from a threefold Cartesian product to \texttt{bool}, the 
type containing true and false only.  
The type variables $\sigma$ and $\sigma'$ input to the type constructor 
\texttt{Kripke} represent the abstract state type and the concrete state type. 
Consequently, the middle element of the triples selected by the relation 
\texttt{mod\_trans} is a function of type $\sigma' \Rightarrow \sigma$ 
mapping elements of the refined state to the abstract state.
The expression in quotation marks after the type is again the
infix syntax in Isabelle that allows the definition of mathematical notation
instead of writing \texttt{mod\_trans} in prefix manner. This nicer infix
syntax is already used in the actual definition.
Finally, the arrow \texttt{\ttImp} is the implication of Isabelle's  meta-logic
while $\ttimp$ is the one of the {\it object} logic HOL. 
They are logically equivalent but of different
types: within a HOL formula $P$, for example, as below $\forall x. P \ttimp Q$, only the implication $\ttimp$
can be used.
\begin{ttbox}
 refinement :: (\ttsigma Kripke \tttimes (\ttsigma' \ttfun \ttsigma) \tttimes \ttsigma' Kripke) \ttfun bool ("_ \ttmref{(\_)} _")
  K \ttmeref K' \ttequiv \ttforall s' \ttin states K'. \ttforall s \ttin init K'. 
             s \ttrelstar{\sigma'} s' \ttimp \ttecal(s) \ttin init K \ttand \ttecal(s) \ttrelstar{\sigma} \ttecal(s')
\end{ttbox}
The definition of \texttt{K \ttmeref\, K'} states that for any state $s'$ 
of the refined Kripke structure that can be reached by the state transition
in zero or more steps from an initial state $s$ of the refined Kripke 
structure, the mapping ${\mathcal E}$ from the refined to the abstract 
model's state must preserve this reachability, i.e., the image of
$s$ must also be an initial state and from there the image of $s'$
under ${\mathcal E}$ must be reached with $0$ or $n$ steps.
